---
title: "Reproducible Research: Peer Assessment 1"
author: "rb wiley"
date: "Wednesday, April 15, 2015"
output: html_document
keep_md: true
---


## Loading and preprocessing the data
- Read the data file into database (previously unzipped in working directory)
- Basic EDA: summarize it
```{r Loadcsv}
activity <- read.csv("activity.csv", )

```
- Basic EDA: summarize data file
```{r SummarizeData}
summary(activity)
```

## What is mean total number of steps taken per day?
- group steps by day
- view as a histogram
```{r}
steps_day <- aggregate(steps ~ date, data = activity, FUN = sum)
barplot(steps_day$steps, names.arg = steps_day$date, xlab = "Date", ylab = "# Steps")
```

- count the days
- calculate mean and median of steps by day
```{r CalcMeanMedian}
nrow(steps_day); mean(steps_day$steps)
median(steps_day$steps)
```
## What is the average daily activity pattern?
- view as time series of average daily steps (y-axis) by 5-in interval (x axis)

```{r PlotIntervalMeans}
interval_mean <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(interval_mean, type = "l")
```

- Which interval has the maximum average # of steps
```{r FindMaxInterval}
interval_mean$interval[which.max(interval_mean$steps)]
```

## Imputing missing values
- Tally the missing values ('NA') in steps variable
- Confirm that tally from summary() function

```{r TallyNAs}
sum(is.na(activity))
##  confirm the tally:
summary(is.na(activity$steps))
```

- replace 'NA' values with interval means
- summarize the new dataset
- to see if replaced 'NAs' changed usable days, compare the counts of days
  - aggregate new daily steps
```{r ImputReplaceNAs}
activity_noNA1 <- merge(activity, interval_mean, by = "interval", 
    suffixes = c("",".ivl"))
NA_cases <- is.na(activity_noNA1$steps)
activity_noNA1$steps[NA_cases] <- activity_noNA1$steps.ivl[NA_cases]
activity_noNA2 <- activity_noNA1[, c(1:3)]
summary(activity_noNA2)
steps_day2 <- aggregate(steps ~ date, data = activity_noNA2, FUN = sum)

```

- count the raw & raw + imputed days; show difference
```{r CountDays} 
no_Raw_days <- nrow(steps_day) 
no_Imputed_days <- nrow(steps_day2) 
```

### `r no_Imputed_days - no_Raw_days` *all 'NA'* days recovered, now included: 
### Raw: `r no_Raw_days` days
### Imputed: `r no_Imputed_days` days


- view histogram of new dataset having imputed missing data
```{r HistoImputedNAs}
steps_day2 <- aggregate(steps ~ date, data = activity_noNA2, FUN = sum)
barplot(steps_day2$steps, names.arg = steps_day2$date, xlab = "Date", ylab = "# Steps")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r WeekdayWeekendComparison}

```
